OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-03-05 paper/paper.Rmd
minor polish

The mothur algorithms produced higher quality OTUs than vsearch in open
reference, closed reference, or _de novo_ modes. However, vsearch was able to
fit more sequence into OTUs than OptiFit in closed reference mode. While both
mothur and vsearch use a dissimilarity threshold for determining how to assign
sequences into OTUs, vsearch is more permissive than mothur. Mothur's OptiFit
and OptiClust use all of the sequences to define an OTU, requiring that all
pairs of sequences (including reference and query sequences) in an OTU are
within the distance threshold without penalizing the MCC. In contrast,
vsearch only requires each query sequence to be similar to the single sequence
that seeded the OTU. In this way, vsearch sacrifices OTU quality in order to
allow more sequences to fit to OTUs. Users who require closed reference
clustering to the greengenes database may prefer to use vsearch if they wish to
maximize the fraction of sequences that can be fit at the cost of OTU quality.
However, mothur's OptiClust or OptiFit are recommended for _de novo_ or open
reference clustering to produce OTUs of the highest possible quality.

When fitting with the split dataset strategy, OTU quality was remarkably similar
when reference sequences were selected by a simple random sample or weighted by
abundance, but quality was slightly worse when sequences were weighted by
similarity. We recommend using a simple random sample since the more
sophisticated reference selection methods do not offer any benefit. The
similarity in OTU quality between OptiClust and OptiFit with this strategy
demonstrates the suitability of using OptiFit to fit sequences to existing OTUs,
such as when using already-trained machine learning models to make predictions
on new data or comparing OTUs across studies. However, when stable OTUs are not
required, we recommend using OptiClust for _de novo_ clustering over the split
strategy with OptiFit since OptiClust simpler to execute but performs similarly.

>>>>

The mothur algorithms produced higher quality OTUs than vsearch in open
reference, closed reference, or _de novo_ modes. However, vsearch was able to
fit more sequence into OTUs than OptiFit in closed reference mode. While both
mothur and vsearch use a distance or similarity threshold for determining how to
assign sequences to OTUs, vsearch is more permissive than mothur. Mothur's
OptiFit and OptiClust use all of the sequences to define an OTU, requiring that
all pairs of sequences (including reference and query sequences) in an OTU are
within the distance threshold without penalizing the MCC. In contrast, vsearch
only requires each query sequence to be similar to the single sequence that
seeded the OTU. In this way, vsearch sacrifices OTU quality in order to allow
more sequences to fit to OTUs. Users who require closed reference clustering to
the greengenes database may prefer to use vsearch if they wish to maximize the
fraction of sequences that can be fit at the cost of OTU quality. However,
mothur's OptiClust or OptiFit are recommended for _de novo_ or open reference
clustering to produce OTUs of the highest possible quality.

When fitting with the split dataset strategy, OTU quality was remarkably similar
when reference sequences were selected by a simple random sample or weighted by
abundance, but quality was slightly worse when sequences were weighted by
similarity. We recommend using a simple random sample since the more
sophisticated reference selection methods do not offer any benefit. The
similarity in OTU quality between OptiClust and OptiFit with this strategy
demonstrates the suitability of using OptiFit to fit sequences to existing OTUs,
such as when using already-trained machine learning models to make predictions
on new data or comparing OTUs across studies. However, when stable OTUs are not
required, we recommend using OptiClust for _de novo_ clustering over the split
strategy with OptiFit since OptiClust is simpler to execute but performs
similarly in terms of both run time and OTU quality.
